### 2020-12-06 Thesis-Chapter2-Background.tex
Address some TODOs

\paragraph*{Volatile Politics} Finally, \ac{the-cw2} is taking place amongst a political landscape marked by
uncertainty, extremism, misinformation, and a pandemic that will have unknown long-term impacts. Since legislature is
part of the broader data security threat model, its behavior is important. Unfortunately, the dynamics of present
leadership and the lack of reliable data on what law enforcement agencies need make it difficult for the technical
community to engage productively with policymakers \cite{granick_2018}.

That brings us to where the debate stands today. A December 2019 U.S. Senate Judiciary Committee featured the top two
U.S. parties both expressing anger over whether data should be ``beyond the reach of the law'' and expressed threats
such as ``get your act together, or we will gladly get your act together for you'' \cite{geller_2019}. In March 2020,
the Republican atop the Senate Judiciary Committee introduced S.3398, the \ac{EARNITAct} of 2020, which seeks to
establish a 16-chair committee with the power to revoke online platforms' liability protection for user-submitted
content if the committee determines the platform is not doing enough to fight \ac{CSAM}, a legal synonym for child
pornography \cite{graham_s3398_2020}. As illustrated by the previously mentioned \textit{New York Times} feature
\cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
largely see the \ac{EARNITAct} as an attempt to indirectly yet effectively outlaw \ac{E2EE} \cite{newman_2020}
\cite{pfefferkorn_2020}.

% NOTE: keep up to date, including the section above mentioning the pandemic
% IDEA: create timeline! https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline

@@

% IDEA: Create an appendix to go into greater detail?
% Could go into these, as well as the commented-out ones below
%  Wiretap Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_119.html)
%  Stored Communications Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_121.html)
%  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)
% TODO: I _really_ need to understand these and how they're used today.

@@

Amended the Wiretap Act, Stored Communications Act, and Pen Register Act. Provided explicit protections for private
% TODO: Briefly introduce these. Even just "Amended the switch-board-era"...
electronic communications data, and updated definitions to contemporary technology. It granted law enforcement explicit
access to certain data as well; some terminology has aged poorly with technological evolution, making the access
increasingly permissive \cite{shamsi_2011}.

@@

% TODO: as of 7/10/20, this bill is still active. Follow up.

@@

    \law{2020}{The \acf{EARNITAct}}{\cite{graham_s3398_2020}}{

Introduced by the chair of the Senate Judiciary Committee, one of the outspoken encryption critics during the 2019
Senate hearing on encryption and \ac{EA}, the \ac{EARNITAct} is designed to address \ac{CSAM} and has received attention
as a threat to encryption \cite{newman_2020} \cite{pfefferkorn_2020}. As described in \mysec{sec-history-current}, the
bill would revoke online platforms' liability protection for user-submitted content if it is does not doing enough to
combat \ac{CSAM}. Critics fear that meeting the regulatory requirements would preclude \ac{E2EE}, and at least one
popular \ac{E2EE} messaging platform has threatened to pull out of the U.S. market if the bill were to pass
\cite{lund_2020}.

% TODO: Update based on new amendment.

@@

Introduced by the same senator as the \ac{EARNITAct}, the \ac{LAEDAct} is a direct attack on encryption. The
\ac{LAEDAct} specifically addresses \ac{DAR} and \ac{DIM} by forcing device, operating system, and application
developers to implement the ability to decrypt all data stored or passing through the device or software system
``concurrently with their transmission'' ``unless the independent actions of an unaffiliated entity make it technically
impossible to do so.'' The act neither mandates nor recommends any technical approach to providing this capability.

% TODO: how does this differ from the old "Compliance with Court Orders Act" draft?

>>>>

\paragraph*{Volatile Politics} Finally, \ac{the-cw2} is taking place amongst a political landscape marked by
uncertainty, extremism, misinformation, and a pandemic that will have unknown long-term impacts. Since legislature is
part of the broader data security threat model, its behavior is important. Unfortunately, the present political dynamics
and the lack of reliable data on what law enforcement agencies need make it difficult for the technical community to
engage productively with policymakers \cite{granick_2018}.

That brings us to where the debate stands today. A December 2019 U.S. Senate Judiciary Committee featured the top two
U.S. parties both expressing anger over whether data should be ``beyond the reach of the law'' and expressed threats
such as ``get your act together, or we will gladly get your act together for you'' \cite{geller_2019}. In March 2020,
the Republican atop the Senate Judiciary Committee introduced S.3398, the \ac{EARNITAct} of 2020, which seeks to
establish a 16-chair committee with the power to revoke online platforms' liability protection for user-submitted
content if the committee determines the platform is not doing enough to fight \ac{CSAM}, a legal synonym for child
pornography \cite{graham_s3398_2020}. As illustrated by the previously mentioned \textit{New York Times} feature
\cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
largely saw the \ac{EARNITAct} as an attempt to indirectly yet effectively outlaw \ac{E2EE} \cite{newman_2020}
\cite{pfefferkorn_2020}. In response, legislators narrowed the focus of the \ac{EARNITAct}, but only after unveiling the
\ac{LAEDAct}

% IDEA: create timeline https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline

@@

% IDEA: Create an appendix to go into greater detail?
% Could go into these, as well as the commented-out ones below
%  Wiretap Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_119.html)
%  Stored Communications Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_121.html)
%  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)
% It would be nice to really understand these and how they're used today. committee_decrypting_2018 describes them.

@@

Included the Stored Communications Act and Pen Register Act, amending the switchboard-era Wiretap Act. Provided explicit
protections for private electronic communications data, and updated definitions to contemporary technology. It granted
law enforcement explicit access to certain data as well; some terminology has aged poorly with technological evolution,
making the access increasingly permissive \cite{shamsi_2011}.

@@

@@

%     \law{2020}{The \acf{EARNITAct}}{\cite{graham_s3398_2020}}{

% Introduced by the chair of the Senate Judiciary Committee, one of the outspoken encryption critics during the 2019
% Senate hearing on encryption and \ac{EA}, the \ac{EARNITAct} is designed to address \ac{CSAM} and has received attention
% as a threat to encryption \cite{newman_2020} \cite{pfefferkorn_2020}. As described in \mysec{sec-history-current}, the
% bill would revoke online platforms' liability protection for user-submitted content if it is does not doing enough to
% combat \ac{CSAM}. Critics fear that meeting the regulatory requirements would preclude \ac{E2EE}, and at least one
% popular \ac{E2EE} messaging platform has threatened to pull out of the U.S. market if the bill were to pass
% \cite{lund_2020}.

% % NOTE: based on the amendment's modifications, it is no longer about encryption. This has become more moderation and
% % CSAM focused, and the LAED Act has taken over the encryption side. A wise split (not that either bill is necessarily
% % good.)
% % https://www.engadget.com/earn-it-act-amendments-pass-senate-judiciary-committee-165030518.html

% }

@@

The \ac{LAEDAct} is a direct attack on encryption and the direct successor of the Compliance with Court Orders Act. It
was introduced by the chair of the Senate Judiciary Committee, also sponsor of the \ac{EARNITAct} and one of the
outspoken encryption critics during the 2019 Senate hearing on encryption and \ac{EA}. The \ac{LAEDAct} specifically
addresses \ac{DAR} and \ac{DIM} by forcing device, operating system, and application developers to implement the ability
to decrypt all data stored or passing through the device or software system ``concurrently with their transmission''
``unless the independent actions of an unaffiliated entity make it technically impossible to do so.'' The act neither
mandates nor recommends any technical approach to providing this capability.
